---
title: "Teaching hydrology with airGRteaching"
author: "Pierre Brigode & Olivier Delaigue"
bibliography:
  - V00_airGRteaching_ref.bib
  - airGR_galaxy.bib
output:
  rmarkdown::html_vignette:
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Ex. 0. Teaching hydrology with airGRteaching}
  %\VignetteEncoding{UTF-8}
---
  
  
```{r, warning=FALSE, include=FALSE}
R_bib <- toBibtex(citation())
R_bib <- gsub("@Manual\\{", "@Manual{Rsoftware_man", R_bib)
airGR_bib <- toBibtex(citation("airGR"))
airGR_bib <- gsub("@Article\\{", "@Article{airGR_art", airGR_bib)
airGR_bib <- gsub("@Manual\\{", "@Manual{airGR_man", airGR_bib)
airGRteaching_bib <- toBibtex(citation("airGRteaching"))
airGRteaching_bib <- gsub("@Manual\\{", "@Manual{airGRteaching_man", airGRteaching_bib)
airGRteaching_bib <- gsub("@Article\\{", "@Article{airGRteaching_art", airGRteaching_bib)
airGRdatasets_bib <- toBibtex(citation("airGRdatasets"))
airGRdatasets_bib <- gsub("@Manual\\{", "@Manual{airGRdatasets_man", airGRdatasets_bib)
options(encoding = "UTF-8")
writeLines(text = c(R_bib, airGR_bib, airGRteaching_bib, airGRdatasets_bib), con = "airGR_galaxy.bib")
options(encoding = "native.enc")
```

```{r, include=FALSE}
formatGR      <- '<strong><font color="#0BA6AA">%s</font></strong>'
GR            <- sprintf(formatGR, "GR")
airGR         <- sprintf(formatGR, "airGR")
airGRteaching <- sprintf(formatGR, "airGRteaching")
airGRdatasets <- sprintf(formatGR, "airGRdatasets")
```

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "figure/")
library(airGRteaching)
Sys.setlocale("LC_TIME", "English")
colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
  } else {
    x
  }
}
```


The purpose of this tutorial is to present teaching activities and exercises that can be done in hydrology classes. 



# Data loading

The figures and simulation performed in this document are based on hydro-meteorological data provided by the `r airGRdatasets` package. The following commands load daily temporal series for one particular catchment.

```{r, echo=TRUE, eval=TRUE}
# Package loading
library(airGRdatasets)

# Catchment data loading
data("B222001001", package = "airGRdatasets")

# Structure of the catchment data object
str(B222001001)
```



# Understanding rainfall-runoff modelling


## The role of model components and parameters

Rainfall-runoff models are composed of different elements, e.g. conceptual reservoirs or unit hydrographs, whose behaviour is defined by equations and parameters. While equations are rarely modified by users, parameter values are a key lever that users can activate to taylor the models to their case studies. Understanding the role of model components and parameters is therefore an unavoidable preliminary step to performing hydrological modelling. The `r airGRteaching` package allows such manipulations.

To illustrate the production and the routing parts of hydrological modelling which are present in any model, it is possible to use the different GR models available using `r airGRteaching` and to produce rainfall-runoff transformations considering different model parameter values.

The GR4J model [@perrin_improvement_2003] comprises a production store (X1 parameter, the maximal capacity of the production store [mm]), which determines the actual evapotranspiration and the net rainfall. The routing of net rainfall is determined through two unit hydrographs (X4 parameter, the unit hydrograph time constant [d]) and a routing store (X3 parameter, the one-day ahead maximal capacity [mm]). A final component represents the intercatchment groundwater exchange (X2 parameter, the exchange coefficient [mm/d]).

The following command lines allow to prepare the hydro-meteorological data for use with GR4J.

```{r, echo=TRUE, eval=TRUE}
# Package loading
library(airGRdatasets)

# Catchment data loading
data("B222001001", package = "airGRdatasets")

# Observed daily time series
ts_obs <- B222001001$TS

# Data processing for GR4J (without Q)
prep_no_q <- PrepGR(DatesR     = ts_obs$Date,
                    Precip     = ts_obs$Ptot,
                    PotEvap    = ts_obs$Evap,
                    HydroModel = "GR4J", 
                    CemaNeige  = FALSE)

# Simulation period
per_sim <- range(prep_no_q$InputsModel$DatesR)
```

As an example, the following commands and figure illustrate the role of the X2 parameter of GR4J on the production part of the rainfall-runoff transformation, by testing different X2 values. We observe higher streamflow values simulated with higher X2 values. 

```{r, v00_fig_gr4j_x2, echo=TRUE, eval=TRUE, warning=FALSE, fig.width=3*3, fig.height=3*1.7, out.width='98%'}
# Different X2 values around its median values (0 [mm/d])
param_x2 <- seq(from = -2, to = 2, by = 1)

# Combination of parameter values (X1, X3 and X4 are fixed; X2 changes)
param_gr4j <- expand.grid(X1 = 350, 
                          X2 = param_x2,
                          X3 = 90,
                          X4 = 1.4)

# Streamflow simulations using parameter sets
sim_x2 <- apply(param_gr4j, MARGIN = 1, FUN = function(i_param_gr4j) {
  i_sim <- SimGR(PrepGR  = prep_no_q, 
                 Param   = i_param_gr4j, 
                 SimPer  = per_sim, 
                 verbose = FALSE)
  i_sim$OutputsModel$Qsim
})

# Graphical comparison
ind_zoom <- 400:430
col_param_x2 <- colorRampPalette(c("green1", "green4"))(ncol(sim_x2))
matplot(x = as.POSIXct(prep_no_q$InputsModel$DatesR[ind_zoom]),
        y = sim_x2[ind_zoom, ],
        xlab = "time [d]", ylab = "flow [mm/d]",
        type = "l", lty = 1, lwd = 2, col = col_param_x2)
legend("topright",
       legend = sprintf("% .1f", param_x2),
       lwd = 2, col = col_param_x2,
       title = "X2 values [mm/d]",
       bg = "white")
```

Moreover, the following commands and figure illustrate the role of the X4 parameter on the routing part of the rainfall-runoff transformation, with delayed flood peak values when considering higher X4 values.

```{r, v00_fig_gr4j_x4, echo=TRUE, eval=TRUE, warning=FALSE, fig.width=3*3, fig.height=3*1.7, out.width='98%'}
# Different X4 values around its median values (1.4 [d])
param_x4 <- seq(from = 1.0, to = 3.0, by = 0.5)

# Combination of parameter values (X1, X2 and X3 are fixed; X4 changes)
param_gr4j <- expand.grid(X1 = 350, 
                          X2 = 0,
                          X3 = 90,
                          X4 = param_x4)

# Streamflow simulations using parameter sets
sim_x4 <- apply(param_gr4j, MARGIN = 1, FUN = function(i_param_gr4j) {
  i_sim <- SimGR(PrepGR = prep_no_q, 
                 Param = i_param_gr4j, 
                 SimPer = per_sim, 
                 verbose = FALSE)
  i_sim$OutputsModel$Qsim
})

# Graphical comparison
ind_zoom <- 400:430
col_param_x4 <- colorRampPalette(c("steelblue1", "steelblue4"))(ncol(sim_x4))
matplot(x = as.POSIXct(prep_no_q$InputsModel$DatesR[ind_zoom]),
        y = sim_x4[ind_zoom, ],
        xlab = "time [d]", ylab = "flow [mm/d]",
        type = "l", lty = 1, lwd = 2, col = col_param_x4)
legend("topright",
       legend = sprintf("% .1f", param_x4),
       lwd = 2,col = col_param_x4, 
       title = "X4 values [d]",
       bg = "white")
```

The relative importance of the production and routing functions depends on the time step considered for the rainfall-runoff simulation. The production processes are more important for the larger time steps (e.g. month or year) since it controls the catchment water balance. This can be easily illustrated by aggregating at an annual scale simulations previously performed at a daily time step. The following commands and figure compare, at an annual scale, the GR4J daily simulations previously performed with different X2 parameter values with the simulations performed with different X4 parameter sets. We can observe that at this annual scale, the impact of considering different X4 parameter values is limited compared to the use of different X2 parameter values.

```{r, v00_fig_sim_aggreg, echo=TRUE, eval=TRUE, fig.width=3*3, fig.height=3*1.7, out.width='98%'}
# Aggregation of the simulated streamflow at the yearly time step
sim_x2_y <- cbind(DatesR = as.POSIXct(prep_no_q$InputsModel$DatesR), 
                  as.data.frame(sim_x2))
sim_x2_y <- SeriesAggreg(x = sim_x2_y, 
                         Format = "%Y",
                         ConvertFun = rep("sum", ncol(sim_x2_y) - 1))
sim_x4_y <- cbind(DatesR = as.POSIXct(prep_no_q$InputsModel$DatesR), 
                  as.data.frame(sim_x4))
sim_x4_y <- SeriesAggreg(x = sim_x4_y, 
                         Format = "%Y",
                         ConvertFun = rep("sum", ncol(sim_x4_y) - 1))

# Graphical comparison
matplot(x = sim_x2_y$DatesR, y = sim_x2_y[, -1], 
        type = "l", lty = 1, lwd = 2, col = col_param_x2, 
        xlab = "time [yr]", ylab = "flow [mm/yr]")
matlines(x = sim_x4_y$DatesR, y = sim_x4_y[, -1], 
         type = "l", lty = 1, lwd = 2, col = col_param_x4)
legend("topright", 
       legend = c("X2", "X4"), 
       lwd = 2, col = c(median(col_param_x2), median(col_param_x4)),
       bg = "white")
```

Finally, the simulation sensitivity on one given parameter can be also addressed using the `r airGRteaching` graphical user interface launched by the `ShinyGR()` function. It is very easy by moving the slider of each parameter to change its value, and to see the effect in real time on the hydrograph of the simulated streamflows.


## On the need to perform a model warm-up

The conceptual reservoirs that compose rainfall-runoff models must be assigned with initial levels when a simulation is done. The way initial levels are defined can lead to potentially significant model errors. The most convenient way that modelers use to initialize rainfall-runoff models is to perform a warm-up run of the model in order to limit the impact of this unknown. 

This issue can be illustrated with `r airGRteaching` by considering different warm-up period lengths. The following commands and figure illustrate a portion of the streamflow simulations obtained considering (i) no warm-up period, (ii) a one-month warm-up period, and (iii) a a four-year warm-up period of the GR4J model. 

Note that if `WupPer` is not set by the user, by default `WupPer = NULL`, and in this case it ensures a one-year warm-up using the time steps preceding the `SimPer` (if data are available).

```{r, v00_fig_wup, echo=TRUE, eval=TRUE, fig.width=3*3, fig.height=3*1.7, out.width='98%'}
# Warm-up and simulation periods
per_wup1m <- c("2002-12-01", "2002-12-31")
per_wup1y <- c("2002-01-01", "2002-12-31")
per_sim   <- c("2003-01-01", "2006-12-31")

# Parameter set
param_gr4j <- c(X1 = 350, X32 = 0, X3 = 90, X4 = 1.4)

# Simulation without warm-up period
sim_wup0d <- SimGR(PrepGR = prep_no_q, 
                   Param  = param_gr4j,
                   WupPer = 0L,
                   SimPer = per_sim)

# Simulation with a 1-month warm-up period
sim_wup1m <- SimGR(PrepGR = prep_no_q,
                   Param  = param_gr4j,
                   WupPer = per_wup1m, 
                   SimPer = per_sim)

# Simulation with a 4-year warm-up period
sim_wup1y <- SimGR(PrepGR = prep_no_q, 
                   Param  = param_gr4j,
                   WupPer = per_wup1y, 
                   SimPer = per_sim)

# Graphical comparison
col_wup <- c("orchid", "orange2", "green3")
matplot(x = as.POSIXct(sim_wup0d$OutputsModel$DatesR),
        y = cbind(sim_wup0d$OutputsModel$Qsim,
                  sim_wup1m$OutputsModel$Qsim,
                  sim_wup1y$OutputsModel$Qsim),
        xlab = "time [d]", ylab = "flow [mm/d]", 
        type = "l", lty = 1, lwd = 2, col = col_wup,
        xlim = as.POSIXct(x = c("2003-01-01", "2003-09-01"), tz = "UTC"))
legend("topright", 
       legend = c("no warm-up", "1-month warm-up", "1-year warm-up"), 
       col = col_wup, lwd = 2,
       bg = "white")
```



# Model calibration


## Manual calibration

In the `r airGRteaching` GUI, it is possible to test different parameter sets of the GR rainfall-runoff models and to estimate the performance of each tested parameter set in order to perform a manual calibration. A classical way to do so through the `r airGRteaching` GUI is to pick an objective function in the table showing the criteria values on the right, to activate the "Show previous simulations (Qold)" button, and to modify parameter values little by little until the simulation and the criterion are satisfactory.

Note that the manual calibration of model parameters is facilitated using the `r airGRteaching` graphical user interface launched by the `ShinyGR()` function. 

You can embed the following code in a loop (except the data processing step using `PrepGR()`). At each iteration you test a new parameter set and compute the corresponding criterion. This way you can find the "best" parameter set.

```{r, v00_fig_cal_manu, echo=TRUE, eval=TRUE, warning=FALSE, fig.width=7*1.5, fig.height=5*1.5, out.width='98%'}
# Data processing for GR4J (with Q for calibration)
prep <- PrepGR(DatesR     = ts_obs$Date,
               Precip     = ts_obs$Ptot,
               PotEvap    = ts_obs$Evap, 
               Qobs       = ts_obs$Qmmd,
               HydroModel = "GR4J", 
               CemaNeige  = FALSE)

# Parameter set to test
i_param_gr4j <- c(X1 = 350, X2 = 0, X3 = 90, X4 = 1.4)

# Rainfall-runoff simulation on the calibration period
i_sim_manu <- SimGR(PrepGR  = prep, 
                    Param   = param_gr4j,
                    WupPer  = c("1999-01-01", "2000-12-31"),
                    SimPer  = c("2001-01-01", "2010-12-31"),
                    EffCrit = "NSE",
                    verbose = TRUE)

# Get the criterion value
GetCrit(i_sim_manu)

# Graphical assessment of the calibration performance
plot(i_sim_manu)
```


## Automatic calibration 

Automatic calibration of model parameters is also possible in `r airGRteaching` using the procedure described by @michel_hydrologie_1991 and by considering one objective function such as NSE [@nash_river_1970] or KGE [@gupta_decomposition_2009]. To do so, there are two options in airGRteaching:

1. clicking on the "Automatic calibration button" in the `r airGRteaching` GUI;
1. using the simple `r airGRteaching` command line functions (`PrepGR()` and `CalGR()`); see following commands.


```{r, v00_fig_cal_auto, echo=TRUE, eval=TRUE, fig.width=7*1.5, fig.height=5*1.5, out.width='98%'}
# Calibration using NSE score 
cal_auto <- CalGR(PrepGR  = prep, 
                  CalCrit = "NSE",
                  WupPer  = c("1999-01-01", "2000-12-31"), 
                  CalPer  = c("2001-01-01", "2010-12-31"))

# Get parameter and criteria values at the end of the calibration step
GetParam(cal_auto)
GetCrit(cal_auto)

# Graphical assessment of the calibration performance
plot(cal_auto)
```


## How to evaluate model calibration? 

Different ways of evaluating the model calibration performance may be conceived using `r airGRteaching`: evaluating criteria on the calibration period, looking at the graphical summary of the calibration performance (displaying using the `plot()` function), comparing simulated and observed streamflows temporal series, etc.

Analyzing simulated versus observed streamflow regimes is an informative indicator of model performance. The following commands and figure compare regimes on a mountainous catchment (located in the French Alps), while the streamflow simulation has been obtained with and without taking into account snow accumulation and melting. The regime comparison might be compelling for the students, hopefully leading them towards the guess of the need to use an additional snow accumulation and melting routine such as CemaNeige [@valery_as_2014], available using `r airGRteaching`.

```{r, v00_fig_cal_eval, echo=TRUE, eval=TRUE, warning=FALSE, fig.width=3*3, fig.height=3*1.7, out.width='98%'}
# Catchment data loading
data("X031001001", package = "airGRdatasets")

# Observed daily time series
ts_obs <- X031001001$TS

# Catchment elevation distribution
hypso <- X031001001$Hypso

# Temporal subset
is_per <- ts_obs$Date >= as.POSIXct("1999-01-01", tz = "UTC") & 
  ts_obs$Date <= as.POSIXct("2009-12-30", tz = "UTC")
ts_obs <- ts_obs[is_per, ]

# Data processing for GR4J (without snow module)
prep_snow_n <- PrepGR(DatesR     = ts_obs$Date,
                      Precip     = ts_obs$Ptot,
                      PotEvap    = ts_obs$Evap, 
                      Qobs       = ts_obs$Qmmd,
                      HydroModel = "GR4J", 
                      CemaNeige  = FALSE)

# Data processing for GR4J with snow module
prep_snow_y <- PrepGR(DatesR     = ts_obs$Date,
                      Precip     = ts_obs$Ptot,
                      PotEvap    = ts_obs$Evap, 
                      Qobs       = ts_obs$Qmmd,
                      TempMean   = ts_obs$Temp,
                      ZInputs    = median(hypso),
                      HypsoData  = hypso,
                      NLayers    = 5,
                      HydroModel = "GR4J", 
                      CemaNeige  = TRUE)

# Calibration using NSE score (without snow module)
cal_snow_n <- CalGR(PrepGR  = prep_snow_n, 
                    CalCrit = "NSE",
                    WupPer  = c("1999-01-01", "2000-12-31"), 
                    CalPer  = c("2001-01-01", "2009-12-30"),
                    verbose = TRUE)

# Calibration using NSE score (with snow module)
cal_snow_y <- CalGR(PrepGR  = prep_snow_y, 
                    CalCrit = "NSE",
                    WupPer  = c("1999-01-01", "2000-12-31"), 
                    CalPer  = c("2001-01-01", "2009-12-30"),
                    verbose = TRUE)

# Combination of observed and simulated streamflow
tab_cal <- data.frame(Date        = cal_snow_n$OutputsModel$DatesR,
                      QOobs       = cal_snow_n$Qobs,
                      Qsim_snow_n = cal_snow_n$OutputsModel$Qsim,
                      Qsim_snow_y = cal_snow_y$OutputsModel$Qsim)

# Computation of regime streamflow
tab_cal_reg <- SeriesAggreg(tab_cal, 
                            Format = "%m", 
                            ConvertFun = rep("mean", ncol(tab_cal) - 1))

# Graphical comparison between simulated and observed streamflow regimes
col_snow <- c("black", rep("orangered", 2))
lty_snow <- c(1, 1:2)
matplot(y = tab_cal_reg[, grep("^Q", colnames(tab_cal))],
        xlab = "time [months]", ylab = "flow [mm/d]", 
        type = "l", lty = lty_snow, lwd = 2, col = col_snow)
legend("topright", 
       legend = c("Qobs", "Qsim without snow mod.", "Qsim with snow mod."),
       lty = lty_snow, lwd = 2, col = col_snow,
       bg = "white")
```


## Streamflow transformation for model calibration

@oudin_dynamic_2006 and many other authors showed the impact of using streamflow transformation during the model calibration. Using `r airGRteaching`, it is possible to apply different streamflow transformations during the model parameter calibration. The following commands and figure compare the simulations performed considering GR4J parameter sets obtained after a calibration on (i) NSE calculated on natural streamflows (noted $NSE_Q$ hereafter), (ii) NSE calculated on square root transformed streamflows (noted $NSE_{\sqrt{Q}}$ hereafter) and (iii) NSE calculated on logarithmic transformed streamflows (noted $NSE_{\log{Q}}$ hereafter), emphazing performance on high, mean and low flows, respectively. 

```{r, v00_fig_flow_transfo, echo=TRUE, eval=TRUE, warning=FALSE, fig.width=3*3, fig.height=3*1.7, out.width='98%'}
# Catchment data loading
data("B222001001", package = "airGRdatasets")
ts_obs <- B222001001$TS

# Data processing for GR4J (with Q for calibration)
prep <- PrepGR(DatesR     = ts_obs$Date,
               Precip     = ts_obs$Ptot,
               PotEvap    = ts_obs$Evap, 
               Qobs       = ts_obs$Qmmd,
               HydroModel = "GR4J", 
               CemaNeige  = FALSE)

# Calibration using NSE score on raw Q
cal_raw <- CalGR(PrepGR  = prep, 
                 CalCrit = "NSE",
                 transfo = "",
                 WupPer  = c("1999-01-01", "2001-12-31"), 
                 CalPer  = c("2002-01-01", "2016-12-31"))

# Calibration using NSE score on sqrt(Q)
cal_sqrt <- CalGR(PrepGR  = prep, 
                  CalCrit = "NSE",
                  transfo = "sqrt",
                  WupPer  = c("1999-01-01", "2001-12-31"),
                  CalPer  = c("2002-01-01", "2016-12-31"))

# Calibration using NSE score on log(Q)
cal_log <- CalGR(PrepGR  = prep, 
                 CalCrit = "NSE",
                 transfo = "log",
                 WupPer  = c("1999-01-01", "2001-12-31"), 
                 CalPer  = c("2002-01-01", "2016-12-31"))

# Combination of simulated streamflow
tab_sim_trsf <- data.frame(Date       = cal_raw$OutputsModel$DatesR,
                           QSIM_rawQ  = cal_raw$OutputsModel$Qsim,
                           QSIM_sqrtQ = cal_sqrt$OutputsModel$Qsim,
                           QSIM_logQ  = cal_log$OutputsModel$Qsim)
tab_sim_trsf <- merge(x = ts_obs[, c("Date", "Qmmd")], 
                      y = tab_sim_trsf, 
                      by = "Date", 
                      all.y = TRUE)

# Computation of regime streamflow
tab_sim_reg <- SeriesAggreg(tab_sim_trsf, 
                            Format = "%m", 
                            ConvertFun = rep("mean", ncol(tab_sim_trsf) - 1))

# Graphical comparison between simulated and observed streamflow regimes
col_trsf <- c("black", rep("orangered", 3))
lty_trsf <- c(1, 1:3)
matplot(y = tab_sim_reg[, -1],
        xlab = "time [months]", ylab = "flow [mm/d]", 
        type = "l", lty = lty_trsf, lwd = 2, col = col_trsf)
legend("bottomleft", 
       legend = c("Qobs", "Qsim", "sqrt(Qsim)", "log(Qsim)"),
       lty = lty_trsf, lwd = 2, col = col_trsf,
       bg = "white")
```


## Impact of objective functions

Similarly to the use of different streamflow transformations during model calibration, the `r airGRteaching` `CalGR()` function allows to test several objective functions such as NSE or KGE. The following commands and figure show how to calibrate GR4J parameters considering two objective functions, and to compare the different simulations on a specific period.

```{r, v00_fig_obj_fun, echo=TRUE, eval=TRUE, warning=FALSE, fig.width=3*3, fig.height=3*1.7, out.width='98%'}
# Calibration using NSE score on Q
cal_nse <- CalGR(PrepGR  = prep, 
                 CalCrit = "NSE",
                 transfo = "",
                 WupPer  = c("1999-01-01", "2001-12-31"), 
                 CalPer  = c("2002-01-01", "2016-12-31"))

# Calibration using KGE score on Q
cal_kge <- CalGR(PrepGR  = prep, 
                 CalCrit = "KGE",
                 transfo = "",
                 WupPer  = c("1999-01-01", "2001-12-31"), 
                 CalPer  = c("2002-01-01", "2016-12-31"))

# Combination of observed and simulated streamflow
tab_crit <- data.frame(Date     = as.POSIXct(cal_nse$OutputsModel$DatesR),
                       Qobs     = cal_nse$Qobs,
                       Qsim_nse = cal_nse$OutputsModel$Qsim,
                       Qsim_kge = cal_kge$OutputsModel$Qsim)

# Graphical comparison
col_crit <- c("black", rep("orangered", 2))
lty_crit <- c(1, 1:2)
matplot(x = tab_crit$Date, y = tab_crit[, -1],
        xlab = "time [d]", ylab = "flow [mm/d]", 
        type = "l", lty = lty_crit, lwd = 2, col = col_crit,
        xlim = as.POSIXct(x = c("2004-01-01", "2004-03-01"), tz = "UTC"))
legend("topleft", 
       legend = c("Qobs", "Qsim NSE", "Qsim KGE"), 
       lty = lty_crit, lwd = 2, col = col_crit,
       bg = "white")
```



# Model evaluation and robustness


## Split-sample test

Split-sample tests, i.e. calibrating a model on a period and evaluating it on a different period [@klemes_operational_1986], is key to assess the transferability of a model, which is most likely used on an extrapolation mode. Using both `CalGR()` and `SimGR()` `r airGRteaching` functions, it is possible to perform split-sample tests for model calibration and evaluation. The following commands and figure compare the performance obtained by the GR4J model after calibration on two sub-periods (`per1` and `per2`) and evaluation on the same sub-periods.

```{r, v00_fig_sst, echo=TRUE, eval=TRUE, warning=FALSE, fig.width=3*3, fig.height=3*1.7, out.width='98%'}
# Calibration and evaluation sub-periods
per1_wup <- c("1999-01-01", "2001-12-31")
per1_sim <- c("2002-01-01", "2008-12-31")
per2_wup <- c("2009-01-01", "2011-12-31")
per2_sim <- c("2012-01-01", "2018-12-31")

# Calibration on per1 and per2
cal_per1 <- CalGR(PrepGR  = prep,
                  CalCrit = "KGE",
                  transfo = "",
                  WupPer  = per1_wup,
                  CalPer  = per1_sim,
                  verbose = TRUE)
cal_per2 <- CalGR(PrepGR  = prep,
                  CalCrit = "KGE",
                  transfo = "",
                  WupPer  = per2_wup,
                  CalPer  = per2_sim,
                  verbose = TRUE)

# Get parameter values at the end of the calibration step
param_per1 <- GetParam(cal_per1)
param_per2 <- GetParam(cal_per2)

# Get criteria values at the end of the calibration step
crit_cal_per1 <- GetCrit(cal_per1)
crit_cal_per2 <- GetCrit(cal_per2)

# Evaluation over per1 and per2
eva_per1 <- SimGR(PrepGR  = prep, 
                  Param   = param_per2,
                  WupPer  = per1_wup,
                  SimPer  = per1_sim,
                  EffCrit = "KGE",
                  verbose = TRUE)

eva_per2 <- SimGR(PrepGR  = prep, 
                  Param   = param_per1,
                  WupPer  = per2_wup,
                  SimPer  = per2_sim,
                  EffCrit = "KGE",
                  verbose = TRUE)

# Get criteria values
crit_eva_per1 <- GetCrit(eva_per1)
crit_eva_per2 <- GetCrit(eva_per2)

# Cleveland dot plot of the criteria
dotchart(c(crit_eva_per1, crit_cal_per1, crit_eva_per2, crit_cal_per2),
         labels = c("eva (per1)", "cal (per1)", "eva (per2)", "cal (per2)"),
         groups = rep(1:2, each = 2),
         col = rep(c("darkred", "darkblue"), each = 2), pch = 19, 
         xlab = "KGE [-]")
```


## Differential split-sample test

The differential split-sample test [@klemes_operational_1986], consists in the identification of two climatically-contrasted periods in the available record and performing the split-sample test using these two periods. The following commands and figure allow to select the wettest and the driest hydrological years (based on the aridity index), and show the calibration/evaluation performance of the GR4J model on these two sub-periods.

```{r, v00_fig_dsst, echo=TRUE, eval=TRUE, warning=FALSE, fig.width=3*3, fig.height=3*1.7, out.width='98%'}
# Estimation of annual aridity index (PE/P)
ts_obs_y <- SeriesAggreg(x = ts_obs[, c("Date", "Ptot", "Evap")], 
                         Format = "%Y", 
                         ConvertFun = c("sum", "sum"), 
                         YearFirstMonth = 10)
ts_obs_y$Arid <- ts_obs_y$Evap / ts_obs_y$Ptot

# Identification of wetter and dryer hydrological years
barplot(height = ts_obs_y$Arid, 
        names.arg = format(ts_obs_y$Date, format = "%Y"), 
        xlab = "time [yr]", ylab = "aridity index [-]",
        col = "royalblue")

# Wet and dry periods
per_wet <- c("2016-10-01", "2017-09-30")
per_dry <- c("2000-10-01", "2001-09-30")

# Calibration over the wet and the dry periods
cal_wet <- CalGR(PrepGR  = prep,
                 CalCrit = "KGE",
                 CalPer  = per_wet,
                 verbose = TRUE)
cal_dry <- CalGR(PrepGR  = prep,
                 CalCrit = "KGE",
                 CalPer  = per_dry,
                 verbose = TRUE)

# Get parameter values at the end of the calibration step
param_dry <- GetParam(cal_dry)
param_wet <- GetParam(cal_wet)

# Get criteria values at the end of the calibration step
crit_cal_dry <- GetCrit(cal_dry)
crit_cal_wet <- GetCrit(cal_wet)

# Evaluation over the wet and the dry periods
eva_wet <- SimGR(PrepGR  = prep, 
                 Param   = cal_dry,
                 SimPer  = per_wet,
                 EffCrit = "KGE",
                 verbose = TRUE)
eva_dry <- SimGR(PrepGR  = prep, 
                 Param   = cal_wet,
                 SimPer  = per_dry,
                 EffCrit = "KGE",
                 verbose = TRUE)

# Get criteria values
crit_eva_dry <- GetCrit(eva_dry)
crit_eva_wet <- GetCrit(eva_wet)

# Cleveland dot plot of the criteria
dotchart(c(crit_eva_dry, crit_cal_dry, crit_eva_wet, crit_cal_wet),
         labels =  c("eva (dry)", "cal (dry)", "eva (wet)", "cal (wet)"),
         col = rep(c("darkorange", "deepskyblue3"), each = 2), pch = 19,
         xlab = "KGE [-]")
```



# Other applications

The basic manipulations of the `r airGRteaching` package illustrated above have also been used in more complete hydrological teaching projects, presented in a vignette format, and exploring different hydrological aspects:

1. streamflow reconstruction using GR2M [@mouelhi_stepwise_2006], a monthly rainfall-runoff model ;
2. low flow forecasting using GR6J [@pushpalatha_downward_2011].
3. climate change impact assessment on a mountainous catchment using GR4J [@perrin_improvement_2003] and its snow accumulation and melt routine CemaNeige [@valery_as_2014].



# References


